### 2016-03-17 doc/draft.Rmd
methods work

In these populations with lower diversity and input titer, we maintained greater than 0.85 sensitivity for SNV at 1% frequency or higher. Our sensitivity was considerably lower for variants at or below 0.5% frequency (`r Hiseq`B and C). We observed an average read depth of over 10,000 reads/bp in these samples. Therefore, the drop in sensitivity is most likely due to the 1,000-fold decrease in nucleaic acid concentration compared to the first data set and the fact that library preparation requires a number of sampling steps that may limit detection. During library preparation, a disproportionate amount of the 2.0\% sample was lost compared to the others, which likely accounts for the reduced sensitivity relative to the 1.0\% sample.

@@

We passaged plaque purified WSN33 and PR8 lab strains 3 times in

@@

Point mutants were made in the WSN33 background using the pHW2000 reverse genetics system (cite)  as previously described. Briefly, primers .......... were used in overlap PCR to generate mutations at HA\_T1583G,HA\_G1006T,HA\_G542T,M\_T861G,M\_A541C,NA\_G1168T,NA\_C454T,NP\_A454C,NP\_A1160T,NS\_G227T,NS\_A809G,PA\_T964G,PA\_T237A,PA\_A1358T,PB1\_G599A,PB1\_G1764T,PB1\_T1288A,PB2\_A1854G,PB2\_A440T,PB2\_A1167T.

@@

Seven hundred and fifty nanograms of purified PCR products were sheared to an average size of 300-400 bp using a Covaris S220 focused ultrasonicator at the following settings : Intensity 4, Duty cycle 10%, Burst/second 200, time 80 seconds. Sequencing libraries were prepared from these fragmented products using the NEBNext Ultra DNA library prep kit (NEB E7370L) and NEBNext multiplex oligonucleotides for Illumina (NEB E7600S). We performed size selection during library preparation using Agencourt AMPure XP beads (Beckman Coulter A63881) with a target insert size of ...... according to the manufactures' instructions.

@@

We extracted viral RNA from infected supernatants using  QIAamp viral RNA minikits (Qiagen 5204) according to the manufactures' instructions, and generated cDNA using Superscript III one-step with HiFi platinum Taq (Invitrogen 12574). PCR products were purified using the Thermo Scientific PCR clean-up according to the manufactures' instructions.

For the WSN33, PR8 data set we used segment specific reactions with the following primers : PB2-Forward-JT	(GCAGGTCAATTATATTCAATATGGAAA), PB2-Reverse-JT	(CAAGGTCGTTTTTAAACTATTCGACAC), PB1-Forward-JT (GCAGGCAAACCATTTGAATGG), PB1-Reverse-JT	(CAAGGCATTTTTTCATGAAGGACAAG), PA-Forward-JT	(GCAGGTACTGATTCAAAATGGAAG), PA-Reverse-JT	(CAAGGTACTTTTTTGGACAGTATGG), NA-Forward-JT	(GCAGGAGTTTAAATGAATCCAAACC), NA-Reverse-JT	(CAAGGAGTTTTTTGAACAAACTACTTG), HA-Forward-JT	(GCAGGGGAAAATAAAAACAACCAAAAT), HA-Reverse-JT	(CAAGGGTGTTTTTCCTTATATTTCTGAA), NP-Forward-JT	(GCAGGGTAGATAATCACTCACAG), NP-Reverse-JT	(CAAGGGTATTTTTCTTTAATTGTCGTACT), M-Forward-JT	(GCAGGTAGATATTGAAAGATGAGTC), M-Reverse-JT	(CAAGGTAGTTTTTTACTCCAGCTCT), NS-Forward-JT	(GCAGGGTGACAAAGACATAATG), NS-Reverse-JT	(CAAAGGGTGTTTTTTATTATTAAATAAGCTG), which with the reaction conditions were adapted from [@Anonymous:A25Ka2tY]. Each 25 $\mu$ l reaction included 1 $\mu$ l of each primer (5 $\mu$ M) and 2 $\mu$ l of RNA template. Reaction conditions were 50$^\circ$ C (60 min), 94$^\circ$ C (2 min), followed by 30 cycles of 94$^\circ$ C (30 sec), 54$^\circ$ C (30 sec), and 68$^\circ$ C (3 min). PCR products were pooled to create equimolar mixes of WSN33 and PR8 segments. The WNS33 sample was then diluted into PR8 serially to yield mixtures in which WSN33 made up 5,2.5,1.25,and 0.63% of the population. We processed the samples for sequencing as below, and sequenced pooled libraries on an Illumina Miseq machine using 2X250 paired end reads.

@@

Reads were aligned to either to PR8 (PR8 WSN33 data set) or WSN33 (pateint derived conditions) using Bowtie2 (cite). Aligments were sorted and processed using Picard (......),variants were called using either DeepSNV or Lofreq and filtered using in house python and R scripts. We connected these steps into an analysis pipeline using bpipe. All commands needed to complete the analysis presented in this work are publically available and free for download at .

>>>>

In these populations with lower diversity and input titer, we maintained greater than 0.85 sensitivity for SNV at 1% frequency or higher. Our sensitivity was considerably lower for variants at or below 0.5% frequency (`r Hiseq`B and C). We observed an average read depth of over 10,000 reads/bp in these samples. Therefore, the drop in sensitivity is most likely due to the 1,000-fold decrease in nucleic acid concentration compared to the first data set and the fact that library preparation requires a number of sampling steps that may limit detection. During library preparation, a disproportionate amount of the 2.0\% sample was lost compared to the others, which likely accounts for the reduced sensitivity relative to the 1.0\% sample.

@@

Madin-Darby canine kidney cells were provided by Arnold S. Monto (University of Michigan School of Public Health) and were maintained in Dulbecco's Modified Eagle Medium (DMEM, Invitrogen) with 10% fetal bovine serum (Gibco and HyClone), 25mM HEPES (Invitogen), and 0.1875% bovine serum albumin (Life Technologies) as in (MFE submitted). WSN33 was rescued from transfections using the 8 plasmid reverse genetic system containing the genomic segments for influenza A/WSN33 (pHW181-188) was a kind gift from Rober Webster (St. Jude's Children's Research Hospital) [@Hoffmann:2002ih]. A biological clone of influenza A/Puerto Rico/8/1934 was obtained from ATCC (VR-1469) and was cloned into the pHW2000 reverse genetic system as before [@Pauly:2015eq]

@@

Point mutants were made in the WSN33 background using the pHW2000 reverse genetics system (cite)  as previously described (MFE citation) . Briefly, we used  PCR mutagenesis to generate mutations at HA\_T1583G, HA\_G1006T, HA\_G542T, M\_T861G, M\_A541C, NA\_G1168T, NA\_C454T, NP\_A454C, NP\_A1160T, NS\_G227T, NS\_A809G, PA\_T964G, PA\_T237A, PA\_A1358T, PB1\_G599A, PB1\_G1764T, PB1\_T1288A, PB2\_A1854G, PB2\_A440T, PB2\_A1167T. Mutants were transfected and passaged once in MDCK cells prior to sequencing.

@@

Seven hundred and fifty nanograms of purified PCR products were sheared to an average size of 300-400 bp using a Covaris S220 focused ultrasonicator at the following settings : Intensity 4, Duty cycle 10%, Burst/second 200, time 80 seconds. Sequencing libraries were prepared from these fragmented products using the NEBNext Ultra DNA library prep kit (NEB E7370L) and NEBNext multiplex oligonucleotides for Illumina (NEB E7600S). We performed size selection during library preparation using Agencourt AMPure XP beads (Beckman Coulter A63881) with a target insert size of 300-400bp or 250 bp for sequencing on Illumina Miseq and Hiseq respectively.

@@

We extracted viral RNA from infected supernaturals using  QIAamp viral RNA minikits (Qiagen 5204) according to the manufactures' instructions, and generated cDNA using Superscript III one-step with HiFi platinum Taq (Invitrogen 12574). PCR products were purified using the Thermo Scientific PCR clean-up according to the manufactures' instructions.

#### PR8 WSN33 population 
WSN33 and PR8 isolates were plaque purified and passaged three times in MCDK cells. We varified the sequences of the viruses using Sanger sequencing.  cDNA was generated in segment specific reactions using : PB2-Forward-JT	(GCAGGTCAATTATATTCAATATGGAAA), PB2-Reverse-JT	(CAAGGTCGTTTTTAAACTATTCGACAC), PB1-Forward-JT (GCAGGCAAACCATTTGAATGG), PB1-Reverse-JT	(CAAGGCATTTTTTCATGAAGGACAAG), PA-Forward-JT	(GCAGGTACTGATTCAAAATGGAAG), PA-Reverse-JT	(CAAGGTACTTTTTTGGACAGTATGG), NA-Forward-JT	(GCAGGAGTTTAAATGAATCCAAACC), NA-Reverse-JT	(CAAGGAGTTTTTTGAACAAACTACTTG), HA-Forward-JT	(GCAGGGGAAAATAAAAACAACCAAAAT), HA-Reverse-JT	(CAAGGGTGTTTTTCCTTATATTTCTGAA), NP-Forward-JT	(GCAGGGTAGATAATCACTCACAG), NP-Reverse-JT	(CAAGGGTATTTTTCTTTAATTGTCGTACT), M-Forward-JT	(GCAGGTAGATATTGAAAGATGAGTC), M-Reverse-JT	(CAAGGTAGTTTTTTACTCCAGCTCT), NS-Forward-JT	(GCAGGGTGACAAAGACATAATG), NS-Reverse-JT	(CAAAGGGTGTTTTTTATTATTAAATAAGCTG), which with the reaction conditions were adapted from [@Anonymous:A25Ka2tY]. Each 25 $\mu$ l reaction included 1 $\mu$ l of each primer (5 $\mu$ M) and 2 $\mu$ l of RNA template. Reaction conditions were 50$^\circ$ C (60 min), 94$^\circ$ C (2 min), followed by 30 cycles of 94$^\circ$ C (30 sec), 54$^\circ$ C (30 sec), and 68$^\circ$ C (3 min). PCR products were pooled to create equimolar mixes of WSN33 and PR8 segments. The WNS33 sample was then diluted into PR8 serially to yield mixtures in which WSN33 made up 5,2.5,1.25,and 0.63% of the population. We processed the samples for sequencing as below, and sequenced pooled libraries on an Illumina Miseq machine using 2X250 paired end reads.

@@

Reads were aligned to either to PR8 (PR8 WSN33 data set) or WSN33 (patient derived conditions) using Bowtie2 (cite). Alignments were sorted and processed using Picard (......),variants were called using either DeepSNV or Lofreq and filtered using in house python and R scripts. We connected these steps into an analysis pipeline using bpipe. All commands needed to complete the analysis presented in this work are public ally available and free for download at .
